Bayesian assessment of heterogeneity in psychological phenomena

### 2023-09-07 ms.Rmd
Small fixes

To calculate a heterogeneity interval, we first specify the desired probability limits. For a 90% interval, we use the .05 and .95 percentiles (which define the central 90% of the distribution). Then, we pass those percentiles and the estimated mean and standard deviation (see Table \@ref(tab:tbl-fit-1)) to the normal quantile function ($q$; `qnorm()` in R) to obtain the interval: $HI_{90} = q([.05, .95], \beta_1, \tau_1) = q([.05, .95],$ `r mu1`, `r sd1`) = [`r hi90`]. In words, this equation calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall within this interval. We illustrate this interval in Figure \@ref(fig:hdist-1)B.

@@

```{r}
#| label: hdist-3
#| fig.height: 2.2
#| fig.width: 4
#| fig-cap: **A**. 100 random draws from the valence effect distribution's posterior distribution. Darker area under the curve represents the proportion of negative effects in the population for that posterior draw ($p^d$). **B**. Histogram of all 4,000 draws from the posterior distribution of $p^d$, and its posterior mean and 95%CI.

@@

```{r}
#| label: hdist-4
#| fig.height: 2.2
#| fig.width: 4
#| fig-cap: **A**. 100 random draws from the valence effect distribution's posterior distribution. Dark area under the curve represents the proportion of effects within a region of practical equivalence to zero (ROPE) in the population for that posterior draw ($p^{ROPE}$). **B**. Histogram of all 4,000 draws from the posterior distribution of $p^{ROPE}$, and its posterior mean and 95%CI.
p1 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) |> 
  arrange(b_valence1) |> 
  # slice(seq(1, nrow(.), by = 2)) |>
  mutate(i = 1:n()) |> 
  mutate(
    dist = dist_normal(b_valence1, sd_person__valence1)
  ) |> 
  ggplot() +
  scale_fill_manual(
    values = c("dodgerblue1", "dodgerblue4"), 
    na.value = "white"
    ) +
  scale_x_continuous(
    "Valence effect",
    expand = expansion(0),
    breaks = extended_breaks()
  ) +
  scale_y_continuous(
    "Sample",
    expand = expansion(0)
  ) +
  aes(xdist = dist, y = i) +
  geom_vline(xintercept = 0, linewidth = .2, linetype = "dashed") +
  stat_slab(
    aes(fill = after_stat(near(x, 0, .1))), 
    height = 15,
    slab_color = "grey10",
    slab_linewidth = .1,
    alpha = .5,
    p_limits = c(.01, .99),
    show.legend = FALSE
  )
p2 <- post1 |> 
  mutate(
    dist = dist_normal(b_valence1, sd_person__valence1),
    pd = cdf(dist, 0),
    prope = cdf(dist, 0.1) - cdf(dist, -0.1)
  ) |> 
  ggplot(aes(prope)) +
  scale_y_continuous(
    "Count",
    expand = expansion(0)
  ) +
  scale_x_continuous(
    TeX('$p^{ROPE}$'),
    breaks = extended_breaks(),
    labels = ~percent(., 1)
  ) +
  stat_histinterval(
    breaks = 50,
    outline_bars = TRUE,
    slab_color = "gray10",
    slab_fill = alpha("dodgerblue1", .33),
    slab_linewidth = .2,
    justification = .01,
    point_color = "dodgerblue4", 
    interval_color = "dodgerblue4",
    .width = .90
  )
(p1 | p2) &
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  ) &
  plot_annotation(tag_levels = "A")
```

@@

```{r}
#| label: hdist-5
#| fig.height: 2.2
#| fig.width: 4
#| fig-cap: **A**. 4,000 random draws from the posterior distribution of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$). **B**. Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its  location $\frac{\tau_1}{\beta_1}$, and its posterior mean and 95%CI.
p1 <- post1 |> 
  ggplot(aes(b_valence1, sd_person__valence1)) +
  scale_y_continuous(
    TeX("$\\tau_1$")
  ) +
  scale_x_continuous(
    TeX("$\\beta_1$")
  ) +
  geom_point(
    shape = 1,
    col = "dodgerblue3",
    alpha = .25,
    size = .5
  )
p2 <- post1 |> 
  mutate(ratio = abs(sd_person__valence1 / b_valence1)) |> 
  ggplot(aes(ratio)) +
    scale_y_continuous(
    "Count",
    expand = expansion(0)
  ) +
  scale_x_continuous(
    TeX('$\\frac{\\tau_1}{\\beta_1}$'),
    breaks = extended_breaks()
  ) +
  stat_histinterval(
    breaks = 50,
    outline_bars = TRUE,
    slab_color = "gray10",
    slab_fill = alpha("dodgerblue1", .33),
    slab_linewidth = .2,
    justification = .01,
    point_color = "dodgerblue4", 
    interval_color = "dodgerblue4",
    .width = .90
  ) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
  )
(p1 | p2) &
  theme(
    aspect.ratio = 1
  ) &
  plot_annotation(tag_levels = "A")
```

>>>>

To calculate a heterogeneity interval, we first specify the desired probability limits. For a 90% interval, we use the .05 and .95 percentiles (which define the central 90% of the distribution). Then, we pass those percentiles and the estimated mean and standard deviation (see Table \@ref(tab:tbl-fit-1)) to the normal quantile function ($q$; `qnorm()` in R) to obtain the interval: $HI_{90} = q([.05, .95], \beta_1, \tau_1) = q([.05, .95],$ `r mu1[1]`, `r sd1[1]`) = [`r hi90`]. In words, this equation calculates the 0.05 and 0.95 quantiles of the normal distribution defined by the mean's ($\beta_1$) and standard deviation's ($\tau_1$) point estimates: We expect 90% of valence effects in the population to fall within this interval. We illustrate this interval in Figure \@ref(fig:hdist-1)B.

@@

```{r}
#| label: hdist-3
#| fig.height: 2.2
#| fig.width: 4
#| fig-cap: "**A**. 100 random draws from the valence effect distribution's posterior distribution. Darker area under the curve represents the proportion of negative effects in the population for that posterior draw ($p^d$). **B**. Histogram of all 4,000 draws from the posterior distribution of $p^d$, and its posterior mean and 95%CI."

@@

```{r}
#| label: hdist-4
#| fig.height: 2.2
#| fig.width: 4
#| fig-cap: "**A**. 100 random draws from the valence effect distribution's posterior distribution. Dark area under the curve represents the proportion of effects within a region of practical equivalence to zero (ROPE) in the population for that posterior draw ($p^{ROPE}$). **B**. Histogram of all 4,000 draws from the posterior distribution of $p^{ROPE}$, and its posterior mean and 95%CI."

p1 <- distinct(dist_samples, Sample, b_valence1, sd_person__valence1) |> 
  arrange(b_valence1) |> 
  # slice(seq(1, nrow(.), by = 2)) |>
  mutate(i = 1:n()) |> 
  mutate(
    dist = dist_normal(b_valence1, sd_person__valence1)
  ) |> 
  ggplot() +
  scale_fill_manual(
    values = c("dodgerblue1", "dodgerblue4"), 
    na.value = "white"
    ) +
  scale_x_continuous(
    "Valence effect",
    expand = expansion(0),
    breaks = extended_breaks()
  ) +
  scale_y_continuous(
    "Sample",
    expand = expansion(0)
  ) +
  aes(xdist = dist, y = i) +
  geom_vline(xintercept = 0, linewidth = .2, linetype = "dashed") +
  stat_slab(
    aes(fill = after_stat(near(x, 0, .1))), 
    height = 15,
    slab_color = "grey10",
    slab_linewidth = .1,
    alpha = .5,
    p_limits = c(.01, .99),
    show.legend = FALSE
  )
p2 <- post1 |> 
  mutate(
    dist = dist_normal(b_valence1, sd_person__valence1),
    pd = cdf(dist, 0),
    prope = cdf(dist, 0.1) - cdf(dist, -0.1)
  ) |> 
  ggplot(aes(prope)) +
  scale_y_continuous(
    "Count",
    expand = expansion(0)
  ) +
  scale_x_continuous(
    TeX('$p^{ROPE}$'),
    breaks = extended_breaks(),
    labels = ~percent(., 1)
  ) +
  stat_histinterval(
    breaks = 50,
    outline_bars = TRUE,
    slab_color = "gray10",
    slab_fill = alpha("dodgerblue1", .33),
    slab_linewidth = .2,
    justification = .01,
    point_color = "dodgerblue4", 
    interval_color = "dodgerblue4",
    .width = .90
  )
(p1 | p2) &
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  ) &
  plot_annotation(tag_levels = "A")
```

@@

```{r}
#| label: hdist-5
#| fig.height: 2.2
#| fig.width: 4
#| fig-cap: Panel **A**. 4,000 random draws from the posterior distribution of the valence effect distribution's mean ($\beta_1$) and standard deviation ($\tau_1$). **B**. Histogram of 4,000 draws from the posterior distribution of the ratio of the valence distribution's scale over its  location $\frac{\tau_1}{\beta_1}$, and its posterior mean and 95\\%CI.

p1 <- post1 |> 
  ggplot(aes(b_valence1, sd_person__valence1)) +
  scale_y_continuous(
    TeX("$\\tau_1$")
  ) +
  scale_x_continuous(
    TeX("$\\beta_1$")
  ) +
  geom_point(
    shape = 1,
    col = "dodgerblue3",
    alpha = .25,
    size = .5
  )
p2 <- post1 |> 
  mutate(ratio = abs(sd_person__valence1 / b_valence1)) |> 
  ggplot(aes(ratio)) +
    scale_y_continuous(
    "Count",
    expand = expansion(0)
  ) +
  scale_x_continuous(
    TeX('$\\frac{\\tau_1}{\\beta_1}$'),
    breaks = extended_breaks()
  ) +
  stat_histinterval(
    breaks = 50,
    outline_bars = TRUE,
    slab_color = "gray10",
    slab_fill = alpha("dodgerblue1", .33),
    slab_linewidth = .2,
    justification = .01,
    point_color = "dodgerblue4", 
    interval_color = "dodgerblue4",
    .width = .90
  ) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
  )
(p1 | p2) &
  theme(
    aspect.ratio = 1
  ) &
  plot_annotation(tag_levels = "A")
```
